---
title: "Most common species in Denmark"
date: "2/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, error = FALSE, warning = FALSE, cache = T)

library(tidyverse)
library(data.table)
library(janitor)
library(DT)

Artlist <- data.table::fread("artsliste.csv") %>% as.data.frame()

library(readr)

frekvens2 <- read_csv("alledata-frekvens2.txt") %>%
  janitor::clean_names() %>%
  as_tibble()

frekvens2$species <- str_split(frekvens2$specieslist, ",")

frekvens2 <- frekvens2 %>%
  dplyr::select(-specieslist) %>%
  unnest(species) %>%
  mutate(species = str_trim(species, side = "both"),
         species = str_remove_all(species, "\\}"),
         site = str_remove_all(site, "\\{"),
         species = as.numeric(species)) %>%
  rename(ArtID = species) %>%
  left_join(Artlist) %>%
  dplyr::select("site", "plot", "year", "ArtID", "NavnDansk",
                "LatArt", "LatFam", "LatOrd", "LatKla",
                "LatDivision", "LatRige") %>% 
  dplyr::filter(LatDivision %in% c("Anthocerophyta", "Bangiophyta", 
"Charophyta", "Chlorophyta", "Cyanophyta", "Magnoliophyta", "Pinophyta", "Pteridophyta", "Xanthophyta"
))  %>% dplyr::mutate(LatArt = str_replace_all(LatArt,"\xeb", "e"),
                      NavnDansk = str_replace_all(NavnDansk,"\\<f8\\>", "\u00f8"),
                      NavnDansk = str_replace_all(NavnDansk,"\\<e6\\>", "\u00e6"),
                      NavnDansk = str_replace_all(NavnDansk,"\\<e5\\>", "\u00e"))
```

# Introduction 

In this document we are characterizing the most common species in Denmark, in order to do that we are using data from NOVANA including `r prettyNum(nrow(frekvens2),big.mark = ".", decimal.mark = ",")` samples in `r prettyNum(length(unique(frekvens2$site)),big.mark = ".", decimal.mark = ",")` sites, and `r prettyNum(length(unique(frekvens2$plot)),big.mark = ".", decimal.mark = ",")` plots sampled from `r min(frekvens2$year)`, till `r max(frekvens2$year)`. This includes `r prettyNum(length(unique(frekvens2$LatArt)),big.mark = ".", decimal.mark = ",")` that have been sampled at least once.

## Number of species per family

The number of species per family in a decreasing order are as follows:

```{r SpeciesPerFamily}
## number of species per family

PerFamily <- frekvens2 %>%
  dplyr::select(LatArt, LatFam) %>%
  distinct() %>%
  group_by(LatFam) %>%
  summarise(Species_number = n()) %>%
  arrange(desc(Species_number)) %>%
  rename(Family = LatFam)

datatable(PerFamily, extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```
## Number of species per Genus

```{r SpeciesPerGenus}
PerGenus <- frekvens2

PerGenus$Genus <- str_split(frekvens2$LatArt," ", simplify = T)[,1] %>% str_trim() 

PerGenus <- PerGenus %>%
  dplyr::select(LatArt, Genus) %>%
  distinct() %>%
  group_by(Genus) %>%
  summarise(Species_number = n()) %>%
  arrange(desc(Species_number))


datatable(PerGenus, extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```

## Number of species per plot

```{r SpeciesPerPlot}
PerSpeciesPlot <- frekvens2 %>%
  dplyr::select("site", "plot", "LatArt", "LatFam", "NavnDansk") %>%
  distinct() %>%
  group_by(LatArt, LatFam, NavnDansk) %>%
  summarise(Ocurrences = n()) %>%
  arrange(desc(Ocurrences)) %>% 
  rename(Family = LatFam,
         Species = LatArt,
         Danish_name = NavnDansk)


datatable(PerSpeciesPlot, extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```

```{r plotlySpecies}

PerSpeciesPlotlyAll <- PerSpeciesPlot %>% 
  ungroup() %>% 
  dplyr::select(Species, Family, Ocurrences) %>% 
  mutate(Family = "All") %>% 
  slice_max(order_by = Ocurrences, n = 100)

TopFamilies <- PerFamily %>% 
  ungroup() %>% 
  slice_max(Species_number, n = 10) %>% 
  pull(Family)

PerSpeciesFamiliesPlotly <- PerSpeciesPlot %>% 
  ungroup() %>% 
  dplyr::select(Species, Family, Ocurrences) %>% 
  dplyr::filter(Family %in% TopFamilies) %>% 
 group_split(Family) %>% 
  purrr::map(~slice_max(.x, order_by = Ocurrences, n = 100)) %>% 
  purrr::reduce(bind_rows) %>% 
  bind_rows(PerSpeciesPlotlyAll)

FamNames <- PerSpeciesFamiliesPlotly$Family %>% unique() %>% sort()

df <- PerSpeciesFamiliesPlotly
```


```{r ThePlot}
library(plotly)

plot_ly() %>%
  add_trace(data = df[df$Family == FamNames[1],] %>% 
              mutate(Species = fct_reorder(Species, desc(Ocurrences))), 
            x = ~Species, 
            y = ~Ocurrences, 
            type = 'bar',
            name = FamNames[1],
            visible = T,
            inherit = F) %>%
  add_trace(data = df[df$Family == FamNames[2],]%>% 
              mutate(Species = fct_reorder(Species, desc(Ocurrences))), 
            x = ~Species, 
            y = ~Ocurrences, 
            type = 'bar', 
            name = FamNames[2],
            visible = F,
            inherit = F) %>%
  add_trace(data = df[df$Family == FamNames[3],] %>% 
              mutate(Species = fct_reorder(Species, desc(Ocurrences))), 
            x = ~Species, 
            y = ~Ocurrences, 
            type = 'bar', 
            name = FamNames[3],
            visible = F,
            inherit = F) %>%
    add_trace(data = df[df$Family == FamNames[4],] %>% 
              mutate(Species = fct_reorder(Species, desc(Ocurrences))), 
            x = ~Species, 
            y = ~Ocurrences, 
            type = 'bar', 
            name = FamNames[4],
            visible = F,
            inherit = F) %>%
    add_trace(data = df[df$Family == FamNames[5],] %>% 
              mutate(Species = fct_reorder(Species, desc(Ocurrences))), 
            x = ~Species, 
            y = ~Ocurrences, 
            type = 'bar', 
            name = FamNames[5],
            visible = F,
            inherit = F) %>%
    add_trace(data = df[df$Family == FamNames[6],] %>% 
              mutate(Species = fct_reorder(Species, desc(Ocurrences))), 
            x = ~Species, 
            y = ~Ocurrences, 
            type = 'bar', 
            name = FamNames[6],
            visible = F,
            inherit = F) %>%
    add_trace(data = df[df$Family == FamNames[7],] %>% 
              mutate(Species = fct_reorder(Species, desc(Ocurrences))), 
            x = ~Species, 
            y = ~Ocurrences, 
            type = 'bar', 
            name = FamNames[7],
            visible = F,
            inherit = F) %>%
    add_trace(data = df[df$Family == FamNames[8],] %>% 
              mutate(Species = fct_reorder(Species, desc(Ocurrences))), 
            x = ~Species, 
            y = ~Ocurrences, 
            type = 'bar', 
            name = FamNames[8],
            visible = F,
            inherit = F) %>%
    add_trace(data = df[df$Family == FamNames[9],] %>% 
              mutate(Species = fct_reorder(Species, desc(Ocurrences))), 
            x = ~Species, 
            y = ~Ocurrences, 
            type = 'bar', 
            name = FamNames[9],
            visible = F,
            inherit = F) %>%
    add_trace(data = df[df$Family == FamNames[10],] %>% 
              mutate(Species = fct_reorder(Species, desc(Ocurrences))), 
            x = ~Species, 
            y = ~Ocurrences, 
            type = 'bar', 
            name = FamNames[10],
            visible = F,
            inherit = F) %>%
    add_trace(data = df[df$Family == FamNames[11],] %>% 
              mutate(Species = fct_reorder(Species, desc(Ocurrences))), 
            x = ~Species, 
            y = ~Ocurrences, 
            type = 'bar', 
            name = FamNames[11],
            visible = F,
            inherit = F) %>%
  layout(
    xaxis = list(
      nticks = 0,
      showticklabels=F
    ),
    updatemenus = list(
      list(y = .7,     
           buttons = list(
             list(method = "update",
                  args = list(list(visible = c(T, F, F, F, F, F, F, F, F, F, F)),
                              list(title = FamNames[1])),
                  label = FamNames[1]),
             list(method = "update",
                  args = list(list(visible = c(F, T, F, F, F, F, F, F, F, F, F)),
                              list(title = FamNames[2])),
                  label = FamNames[2]),
             list(method = "update",
                  args = list(list(visible = c(F, F, T, F, F, F, F, F, F, F, F)),
                              list(title = FamNames[3])),
                  label = FamNames[3]),
             list(method = "update",
                  args = list(list(visible = c(F, F, F, T, F, F, F, F, F, F, F)),
                              list(title = FamNames[4])),
                  label = FamNames[4]),
             list(method = "update",
                  args = list(list(visible = c(F, F, F, F, T, F, F, F, F, F, F)),
                              list(title = FamNames[5])),
                  label = FamNames[5]),
             list(method = "update",
                  args = list(list(visible = c(F, F, F, F, F, T, F, F, F, F, F)),
                              list(title = FamNames[6])),
                  label = FamNames[6]),
             list(method = "update",
                  args = list(list(visible = c(F, F, F, F, F, F, T, F, F, F, F)),
                              list(title = FamNames[7])),
                  label = FamNames[7]),
             list(method = "update",
                  args = list(list(visible = c(F, F, F, F, F, F, F, T, F, F, F)),
                              list(title = FamNames[8])),
                  label = FamNames[8]),
             list(method = "update",
                  args = list(list(visible = c(F, F, F, F, F, F, F, F, T, F, F)),
                              list(title = FamNames[9])),
                  label = FamNames[9]),
             list(method = "update",
                  args = list(list(visible = c(F, F, F, F, F, F, F, F, F, T, F)),
                              list(title = FamNames[10])),
                  label = FamNames[10]),
             list(method = "update",
                  args = list(list(visible = c(F, F, F, F, F, F, F, F, F, F, T)),
                              list(title = FamNames[11])),
                  label = FamNames[11])
           )))
  )

```



## Number of plots per habitat type

```{r HabitatType}
### Model habitat type

Sites <- data.table::fread("/home/derek/Documents/denmark_stratification/Novana/alledata-abiotiske.csv") %>%
  as.data.frame() %>%
  dplyr::select("site", "plot", "year", "UTMx", "UTMy", "sekhabtype") %>%
  mutate(site = as.character(site)) %>%
  left_join(frekvens2)


ForModel <- Sites %>%
  dplyr::select("sekhabtype","LatArt")



ForModel %>% dplyr::group_by(sekhabtype) %>% summarise(n = n()) %>% dplyr::arrange(desc(n)) %>% datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))

```

